Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-06-25 numerical-analysis.tex
Edited Section 10 (analysis).  Planning to skip to sections 12-14, and then return to 11 later.

Detailed analysis of simulation results requires both the tools to ask
sophisticated questions of the data and the ability to process vast quantities of
data at high time-cadence.  As simulations grow in size and complexity, storing data
for post-processing simply becomes intractable.  To cope with this, we have
instrumented \enzo\ with the ability to conduct analysis during the course of a
simulation.  This enables analysis with extremely high time cadence (even as often
as every subcycle of the finest refinement level), without
attempting to write an entire checkpoint output to disk.  The current mechanism
for conducting analysis in \enzo\ during the course of the simulation utilizes
the same compute resources as used by the simulation itself by transferring their usage 
from \enzo\ to the analysis routines; this is often
referred to as \textit{in situ} analysis or visualization.  Utilizing a
dynamically-scheduled second set of computing resources, often referred to as
\textit{co-scheduled} analysis or visualization, provides greater flexibility and overall
throughput at the expense of simplicity.

@@

Where microphysical solvers or other operator-split physics calculations can be
done in Python, \texttt{yt} can serve as a driver for these calculations.  A
major feature set that is currently being developed is to pass structured
(i.e., non-fluid) information back from \texttt{yt} into \enzo.  For instance,
this could be the result of semi-analytic models of the growth and evolution of
star clusters, galaxy particle feedback parameters that have been influenced by
merger-tree analysis, or even spectral energy distributions that are calculated
within \texttt{yt} and provided to \enzo\ as input into radiative transfer
calculations.  Future versions will include this, as well as the ability to
dynamically allocate compute resources to \texttt{yt} such that the simulation
may proceed asynchronously with analysis (\textit{co-scheduled} analysis).
With this functionality will also come the ability to dynamically partition
data, such that spatially-decomposed operations such as volume rendering become
feasible during the course of a simulation.

@@

\enzo\ saves tracer particle data at user-specified intervals,
independent of the intervals at which regular data sets are written.  The data written out
typically includes the tracer particle's unique ID, position, velocity, density,
and temperature, but is easily extensible to output any grid-based
quantity that a user requires.  This capability has been used quite
effectively in several papers, including \citet{2010ApJ...715.1575S}
and \citet{2012ApJ...748...12S}. It is crucial to keep in mind that
tracer particles model a fixed number of Lagrangian fluid trajectories. The fluid
on the grid, however, models the motion of all the mass and represents
the average quantity in a grid cell's volume. Consequently, after a period of
evolution, tracer particles --even if they initially had the same
density distribution as the gas-- will not have the same density
distribution as the fluid. For example, they tend to accumulate at stagnation points
of the flow, and care has to be taken in using these particles
appropriately. Tracer particles are very useful in studies such as the variety of
histories of the hydrodynamic quantities in Lagrangian fluid elements
and when evaluating complex chemical and cooling models in regard to the
simpler ones used in the actual numerical evolution.

@@

where $\myvec{v}$ is the velocity field, $T$ is the temperature, $\rho$ is the
density, and $S=T/\rho^{\gamma-1}$ is the entropy.  $T_2$ and $T_1$ are the post-shock~(downstream)
and pre-shock~(upstream) temperatures, respectively. An optional temperature floor
may be chosen, which is useful for situations such as cosmological simulations without a radiation
background where underdense gas in the intergalactic medium (IGM) can cool adiabatically to
very low temperatures.

@@

Shock finding in the context of AMR is applied grid-by-grid.  If a search for
pre/post-shock cells goes outside the bounds of the grid ghost zones, the
search is stopped for that particular shocked cell. In most situations this is
adequate since the hydrodynamic shock is captured in fewer than the number of
ghost zones.  Shock finding can be run either upon data output, or at each step
in the evolution of an AMR level if the Mach and pre/post-shock quantities are
needed for additional physics modules.

>>>>

Detailed analysis of simulation results requires both the tools to ask
sophisticated questions of the data and the ability to process vast quantities of
data at high time-cadence.  As simulations grow in size and complexity, storing data
for post-processing simply becomes intractable.  To cope with this, we have
instrumented \enzo\ with the ability to conduct analysis during the course of a
simulation.  This enables analysis with extremely high time cadence (as often
as every subcycle of the finest refinement level), without
attempting to write an entire checkpoint output to disk.  The current mechanism
for conducting analysis in \enzo\ during the course of the simulation utilizes
the same computional resources as are used by the simulation itself by transferring their usage 
from \enzo\ to the analysis routines; this is often
referred to as \textit{in situ} analysis or visualization.  Utilizing a
dynamically-scheduled second set of computing resources, often referred to as
\textit{co-scheduled} analysis or visualization, provides greater flexibility and overall
throughput at the expense of simplicity.

@@

Where microphysical solvers or other operator-split physics calculations can be
done in Python, \texttt{yt} can serve as a driver for these calculations.  A
major feature set that is currently being developed is to pass structured
(i.e., non-fluid) information back from \texttt{yt} into \enzo.  For instance,
this could be the result of semi-analytic models of the growth and evolution of
star clusters, galaxy particle feedback parameters that have been influenced by
merger-tree analysis, or even spectral energy distributions that are calculated
within \texttt{yt} and provided to \enzo\ as input into radiative transfer
calculations.  Future versions will include this, as well as the ability to
dynamically allocate computional resources to \texttt{yt} such that the simulation
may proceed asynchronously with analysis (\textit{co-scheduled} analysis).
With this functionality will also come the ability to dynamically partition
data, such that spatially-decomposed operations such as volume rendering become
feasible during the course of a simulation.

@@

\enzo\ saves tracer particle data at user-specified intervals,
independent of the intervals at which regular data sets are written.  The data written out
typically includes the tracer particle's unique ID and position, as
well as the velocity, density,
and temperature of the gas at its location, but is easily extensible to output any grid-based
quantity that a user requires.  This capability has been used quite
effectively in several papers, including \citet{2010ApJ...715.1575S}
and \citet{2012ApJ...748...12S}. It is crucial to keep in mind that
tracer particles model a fixed number of Lagrangian fluid trajectories. The fluid
on the grid, however, models the motion of all the mass and represents
the average quantity in a grid cell's volume. Consequently, after a period of
evolution, tracer particles -- even if they initially had the same
density distribution as the gas -- will not have the same density
distribution as the fluid. For example, they tend to accumulate at stagnation points
of the flow, and care has to be taken in using these particles
appropriately. Tracer particles are very useful in studies such as the variety of
histories of the hydrodynamic quantities in Lagrangian fluid elements
and when evaluating complex chemical and cooling models in regard to the
simpler ones used in the actual numerical evolution.

@@

where $\myvec{v}$ is the velocity field, $T$ is the temperature, $\rho$ is the
density, and $S=T/\rho^{\gamma-1}$ is the entropy.  $T_2$ and $T_1$ are the post-shock~(downstream)
and pre-shock~(upstream) temperatures, respectively. An optional temperature floor
may be chosen, which is useful for situations such as cosmological simulations without a radiation
background where underdense gas in the intergalactic medium (IGM) can cool adiabatically to
unphysically low temperatures.

@@

Shock finding in the context of AMR is applied grid-by-grid.  If a search for
pre/post-shock cells goes outside the bounds of the grid ghost zones, the
search is stopped for that particular shocked cell. In most situations this is
adequate since the hydrodynamic shock is captured in fewer than the number of
ghost zones.  Shock finding can be run either upon data output or at each step
in the evolution of an AMR level if the Mach and pre/post-shock quantities are
needed for additional physics modules.
